---
title: "Penguin Bill Length Analysis (R)"
author: "Edmund Yong"
date: "2026/09/21"
format: 
  html:
    code-fold: true
---

## Goal
The goal of this post is to showcase the varying bill lengths in millimeters for different penguin species by sex.

### Data Source
The data used in this post is sourced from the [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/) dataset shown below in @tbl-palmerpenguins-dataset.

```{r}
#| label: tbl-palmerpenguins-dataset
#| tbl-cap: "Palmer Penguins Dataset"
#| message: false
#| warning: false
library(tidyverse)
library(palmerpenguins)

head(penguins)
```

### Data Cleanup

The cleaned data can be found below in @tbl-penguin-bill-length. To clean the data, records without a recorded sex were removed from the original dataset. The mean of the penguins' bill length was then calculated after grouping by the penguins' species and sex.

```{r}
#| label: tbl-penguin-bill-length
#| tbl-cap: "Bill length of penguins by species (Table)."
#| message: false
#| warning: false
modified_penguins <- penguins
modified_penguins <- drop_na(modified_penguins, sex)
modified_penguins <- modified_penguins |>
  group_by(species, sex) |>
  summarise(bill_length_mm = mean(bill_length_mm))

modified_penguins

```

### Findings
After cleaning the data and viewing the data it in @fig-penguin-bill-length below, it can be observed that male penguins on average across the listed species have longer bills compared to female penguins.

```{r}
#| label: fig-penguin-bill-length
#| fig-cap: "Bill length of penguins by species (Chart)."
#| message: false
#| warning: false
modified_penguins_plot <- ggplot(data = modified_penguins, mapping = aes  (x = species,
    y = bill_length_mm,
    fill = sex)) +
    geom_col(position='dodge') +
    xlab("Species") +
    ylab("Bill Length (mm)") +
    theme_bw() +
    theme(text = element_text(size = 18))
modified_penguins_plot
```
Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/),
Palmer Station Antarctica LTER.
